---
title: "esm206_lab6_matthew_mccafferty"
author: "Matthew McCafferty"
date: "11/2/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
library(broom)
library(equatiomatic)
```

## Rank Based Test example

Mann-Whitney U (non-parametric alternative to an unpaired 2 sample t-test).

H0: Equal ranks (or medians)
HA: Populations where ranks are not equal (or, medians not equal)

### Create Samples

'set.seed()' to create "pseudorandom, but reproducible samples"


```{r}
set.seed(1414)
gp_1 <- sample.int(20, size = 15, replace = TRUE)

set.seed(1424)
gp_2 <- sample.int(30, size = 15, replace = TRUE)
```

Take a quick look at the distribution using 'hist()'

```{r}
hist(gp_1)
hist(gp_2)
```

Try a t-test so we can compare the results:

```{r}
ttest <- t.test(gp_1, gp_2)

# If these data were paired, I coan just add the argument 'paired= "TRUE" within the t.test() function. 
```

This p.value means that if our samples are drawn from populations with the same mean (i.e., if the null hypothesis is true) there is a `r round(ttest$p.value, 2)` propbability of finding sample means *atleast as different* as mine by random chance.  
[A more comprehensive explanation of difference between groups... actually means, data spread, effect size] there is no significant difference (t(df) = t-statistic value, p = ).

### Mann Whitney U rank based test

We use the 'wilcox.test()' function to perform a two-sample rank-based test.

```{r}
mwu <- wilcox.test(gp_1, gp_2)
```

what does that *p*-value of 0.28 actually mean? It means that if the null hypothesis is true (these samples were drawn from populations with the same median), there is a probability of 0.28 that we could have found median values *at least as different as ours* by chance. In other words: not sufficient evidence to reject the null hypothesis of equal ranks (or medians) using a significance level of 0.05.

This p value means that if our samples our drawn from a population with the same median, there is a 28% chance of finding sample medians at leaset as different as ours by random chance.

There is no sigificant difference in medians (Mann-Whitney U; U = `rmwu$statistic`, p = `r round(mwu$p.value, 2)`).

## Simple Linear Regression

Two variables *continuous quantitative* that you are trying to understand the realationship between, and you think that a linear relationship makes sense.

Use the `penguins` data from palmer penguins , exploring the flipper length (x) versus body mass(y)

```{r}
# use view function to look at
# Make ecporatory scatterplot of penguin flipper length in x-axis vers body mass y axis

ggplot(data = penguins, aes(x= flipper_length_mm, y= body_mass_g)) +
  geom_point() +
  geom_smooth(method = "lm")

```
### Find the model estimates (coefficients)

Use the `lm` function for linear model. 

```{r}
penguin_lm <- lm(body_mass_g ~ flipper_length_mm, data = penguins)

summary(penguin_lm)

```
Both the intercept and flipper_length_mm coefficients are significantly different from zero (not super interesting)
- The R^2^ value is 0.759 - meaning that 75.9% of variance in body mass is explained by flipper length
49.69g per mm (flipper length)

intercept -5780.83 is the expected average mass with penguins that have a flipper length of zero

```{r}
extract_eq()
```

Model Coefficients : `r penguin_lm$coefficients`
Flipper length coefficient: `r penguin_lm$coefficients[2]`

### Use `broom:: tidy()` to get a tidy output from lm()

```{r}
penguin_lm_tidy <- tidy(penguin_lm)

penguin_lm_tidy
```
### Use broom::glance() to get output of model fit and statis

```{r}
penguin_lm_glance <- glance(penguin_lm)

penguin_lm_glance
```
### Use equamatic to reurn model to latex
```{r}
extract_eq(model = penguin_lm, use_coefs = TRUE)
```

$\frac{apples}{bananas}$

###Use plot function to automaticall produce diagnostic plots

```{r}
plot(penguin_lm)
```


